### 2012-09-11 draft.tex
added GPU Implementation

\documentclass[times, 10pt, twocolumn]{article}
\usepackage{latex8}
\usepackage[dvips]{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\usepackage{txfonts}
\usepackage{courier}
\usepackage{subfigure}
\usepackage{comment}
\usepackage{url}
\usepackage[square,numbers]{natbib}

\newcommand{\hd}{\emph{H+D}}
\newcommand{\hp}{\emph{Hpin}}
\newcommand{\dm}{\emph{Dmap}}

%-----------------------------------------
% need for camera-ready
%\pagestyle{empty}

%----------------------------------------

\title{Accelerating Traffic Simulation and Its Real-Time Challenges}

\author {
Manato Hirabayashi, Shinpei Kato, Masato Edahiro\\
\textit{Department of Information Engineering}\\
\textit{Nagoya University}\\
\and
Yuki Sugiyama\\
\textit{Department of Physics}\\
\textit{Nagoya University}\\
}

\begin{document}

\maketitle

%-----------------------------------------
% need for camera-ready
%\thispagestyle{empty}

\begin{abstract}
\end{abstract}

\section{Introduction}

Transportation systems underlie our industry and life as part of
societal infrastructure.
Economy depends highly on traffic flow.
Once a traffic jam occurs, for example, it could cause significant
economic damage through degradation of transport efficiency, energy
consumption, and environmental poisoning.
According to the Japanese government report a few years ago, economic
losses due to traffic congestion reach one hundred billion dollars per
year in Japan.
Albeit a major issue of transportation systems, the mechanism of traffic
jam is not well-explained in the literature.
What is particularly important is ``phantom jam'', which often happens
on freeways without any accident.
This phantom jam should be avoided by science, given that it is
a physical phenomenon but not one caused by human errors.
However, we are not aware of scientific and engineering solutions to
such real-world problems.

In physics, traffic flow is described by mathematical
models~\cite{Bando1995, Kerner1993, Nagel1992}.
While their mathematical expressions are not identical, they are all
compute-intensive procedures.
For example, traffic simulation based on the Optimal Velocity (OV)
model~\cite{Bando1995} computes locations and velocities of agents every
sampling period, solving OV equations. 
Specifically, the location $x_n$ of the $n$th agent is described by the
following equation, where $\Delta x_n = x_{n+1} - x_n$ is a distance to
a preceding agent, $a$ is a sensitivity, and $V()$ is an optimal
velocity function:
\begin{eqnarray}
 \label{eqn:ov}
 \frac{d^2 x_n}{d t^2} = a \left\{V(\Delta x_n) - \frac{d x_n}{d t}\right\}.
\end{eqnarray}

Applying a large number of agents to simulation using the above formula,
it is apparent that computational workload increases exponentially.
In fact, the above formula considers only one dimension, which restricts
applications of simulation to freeway traffic flows, powder flows,
molecular motors, and so on.
Making it multi-dimensional further increases computational workload,
while allowing more complicated simulations, such as traffic networks,
internet packet flows, evacuation routes, and herd formations of
animals, to use similar optimal velocity models.
Given a scale of million agents in the real world, traffic
simulation should be supported by powerful computer systems.

Deployment of traffic simulation may require real-time feedback from the
real world.
This turns out to be a cyber-physical systems (CPS) problem.
Real-time traffic simulation is another challenging issue, where
the rate and the preciseness of simulation must be traded to meet the
requirement of a given scenario.
For instance, evacuation may want very high rate simulation, sacrificing
the preciseness of simulation to some extent.
Unfortunately, none of those challenges has been explored yet, largely
due to a lack of multidisciplinary research collaborations between physics
and computer science.

This paper explores how to accelerate traffic simulation using the
state-of-the-art parallel computing technology.
In particular, we use the graphics processing unit (GPU), which
integrates hundreds of cores on a chip.
Recent GPUs are becoming more and more suitable for general-purpose
data-parallel applications.
The traffic simulation program used in this paper applies
Equation~\eqref{eqn:ov} to a large number of agents.
The resulting workload is highly data-parallel and compute-intensive,
which can be nicely offloaded on to the GPU.
We also identify the bottleneck in accelerating our traffic simulation
program using the GPU, and provide some insight into its solution.

\section{Assumption}
\label{sec:assumption}

grid, block, thread.

A unit of functions that is launched on the GPU is called a \textit{kernel}.

\section{Traffic Simulation}
\label{sec:traffic_simulation}

\begin{enumerate}
 \item Initialize the time $t$, and set the initial values of $x_n(t)$
       (also $y_n(t)$ and $z_n(t)$, if necessary for multidimensional
       versions).
 \item Increase the time $t$ by the sampling period $\Delta t$.
 \item Compute the location $x_n(t)$ (also $y_n(t)$ and $z_n(t)$, if
       necessary for multidimensional versions), and the velocity
       $v_x(t)$ at time $t$ for each agent $A_n$, using the OV model.
 \item Go back to Step 2, if the simulation time is expired.
 \item Exit the program.
\end{enumerate}

Note that Step 3 exploits a lot of loop procedures to derive the
locations and velocities of agents.
This part can be accelerated by the GPU.

\section{GPU Implementation}
\label{sec:gpu_implementations}

GPU performance is dominated by the program design.
GPU-accelerated programs are typically divided into two pieces of code.
The CPU code plays a role of a master thread that controls the program
flow.
The GPU code, on the other hand, spawns a bunch of worker threads to
execute compute-intensive parts of the program in parallel, thus
accelerating the overall program.
What is often argued in performance optimization is how to parallelize
the compute-intensive parts into threads.
This is actually the well-studied problem in the literature.
What is not really understood yet is when to offload the program on to
the GPU.
This paper explores two schemes that use the GPU at different timings to
see how GPU performance is affected.

\subsection{Sample-in-CPU Scheme}
\label{sec:sample-in-cpu}

We first implement such a scheme that uses the GPU \textit{only if
necessary}.
In other words, the program is offloaded on to the GPU, only when
computations of locations and velocities of agents are accelerated by
parallelization.
The CPU bridges across sampling periods to manage simulation.
In this scheme, the control flow of simulation is always returned to the
CPU at the end of each sampling period.
Such a synchronized approach makes the programmer easy to
obtain intermediate results of simulation, whereas the overhead imposed
on moving back and forth between the CPU and the GPU must be
compromised.

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-cpu.eps}
\caption{Block diagram of the Sample-in-CPU scheme.}
\label{fig:sample-in-cpu}
\end{figure}

Figure~\ref{fig:sample-in-cpu} shows a brief overview of the
Sample-in-CPU scheme.
This scheme also has several alternatives depending on how many kernels
need to be launched on the GPU in each period. 
Suppose that we want to offload $n$ pieces of \textit{for} loops on to
the GPU.
We may return to the CPU $n$ times in totall, that is, return every time
one \textit{for} loop breaks, or otherwise we may just return once when
all the $n$ \textit{for} loops end.
This is a design decision, and is also dependent on the GPU architecture
and the parallelization structure.
In general, parallelized threads need to be synchronized when moving
across basic blocks.
Specifically, when moving to the next \textit{for} loop executed in
parallel, all the threads relevant to this parallel computing procedure
must synchronize with each other.
However, the maximum number of threads that can be synchronized on the
GPU is often limited.
As of 2012, for example, NVIDIA's GPU architectures limit the number of
such threads to 1024 or less.
Our implementation therefore forces the program to return to the CPU
every time one \textit{for} loop breaks.
Note that this is not a conceptual limitation of GPU computing, but is a
current limitation of hardware.
We believe that this limitation would be removed or mitigated in
next-generation GPU architectures.

\subsection{Sample-in-GPU Scheme}
\label{sec:sample-in-gpu}

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-gpu.eps}
\caption{Block diagram of the Sample-in-GPU scheme.}
\label{fig:sample-in-gpu}
\end{figure}

We next implement such a scheme that uses the GPU \textit{all the time},
even to control simulation.
There is one big kernel running on the GPU, which is launched only once
at the beginning.
After offloading the program on to the GPU, the CPU is going to wait for
the completion of simulation.
This scheme is almost optimized in performance, since there is little
overhead in communication between the CPU and the GPU.
The downside of this scheme is that the CPU and the GPU are not
synchronized.
The progress of simulation is not visible from the CPU, unless the
program implements a specific interface to allow the CPU to access
intermediate results of simulation running on the GPU.
In our implementation, we consider providing a framework that the CPU
downloads data from the GPU asynchronously without awareness of
simulation, when the user requests intermediate results.
This asynchronous data access may affect the performance of simulation,
though in reality this access would not happen more than once in a
sampling period.

Figure~\ref{fig:sample-in-gpu} shows a block diagram of the
Sample-in-GPU scheme.
The procedure is very simple.
The most portion of code of simulation is executed on the GPU.
Given that the single-thread performance of recent GPUs is getting more
reliable, this approach is pretty reasonable.
As mentioned in Section~\ref{sec:sample-in-cpu}, however, the current
limitation of the GPU architecture prevents us from synchronizing among
blocks at a scale of thousands threads.
Therefore, this scheme is speculative in a sense that it does not work today
but may appear in the future.

This paper provides some degree of insights into how this scheme is
effective.
We implement this scheme with the current GPU architecture under the
assumption that global synchronization among blocks works.
There is also another possible approach for this scheme that we limit
the number of threads to what is supported by the GPU architecture.
This alternative implementation, however, is left open for future work.

\section{Evaluation}
\label{sec:evaluation}

We evaluate the performance of our GPU-accelerated traffic simulation
programs, using an NVIDIA GeForce GTX 560 Ti graphics card.
This is a middle-end graphics card integrating 394 compute cores on a
chip.

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_accel.eps}
\caption{Acceleration of traffic simulation.}
\label{fig:eval_accel}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_nosync.eps}
\caption{Impact of synchronization.}
\label{fig:eval_nosync}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_error.eps}
\caption{Error in location and velocity due to imprecise synchronization.}
\label{fig:eval_error}
\end{figure}

\subsection{Experimental Results}
\label{sec:experimental_results}

\subsection{Discussion}
\label{sec:discussion}

Imprecise Computation.
Gdev~\cite{Kato2012}.

\section{Conclusion}
\label{sec:conclusion}

\bibliographystyle{plain}
{\footnotesize
\bibliography{references}
}

\end{document}